Big Ideas in Sports Analytics and Statistical Tools for their Investigation

### 2022-07-16 sports_analytics.Rmd
python

Every sport is different, with its own set of rules, strategies, methods of data collection, number of players, and the magnitude of the role of chance. 
At the same time, many sports are similar, either because one evolved from the other, or the structure of the game shares certain attributes. 
Sports that are closely related anthropologically may or may not share common applications of analytical methods. 
For example, although belonging to the same bat-and-ball family, baseball and cricket differ in strategies such as batting order or sacrifice plays.
Conversely, with just a few small tweaks, analytical metrics might work just as well across sports that are unrelated and quite different.
For instance, an Elo rating could be equally valid for chess players and ice hockey teams.

@@

A more in-depth overview of the history of expected points is provided in @Yurko2019nflwar (Section 1.1). 
Most importantly, this paper uses publicly available data provided by the **nflscrapR** package [@R-nflscrapR] to model the expected points on a play-by-play level in football.
The authors introduce a multinomial logistic regression approach, which takes into account the current down, time remaining, yards from endzone, yards to go, and the indicators for goal to go and whether the time remaining in the half is less than two minutes. 
This model estimates the probabilities of the following possible scoring outcomes after each play: no score, safety, field goal, and touchdown for both the offensive and defensive teams, all of which have a point value. 
The expected points for a play can then be calculated accordingly, by summing up the products of the scoring event point values and their associated probabilities.

@@

The notion of EPV has also been extended to association football (soccer.)
@fernandez2021framework implement deep learning methods to examine the instantaneous expected value of soccer possessions. 
This approach considers passes, ball drives, and shots in soccer as the main set of actions used to compute expected possession value. 
Many applications can be derived from this framework, including predicting which footballer is most likely to receive the next pass from the current on-ball player.

@@

```{r chess, fig.show="hold", out.width="49%", fig.cap="At left, the starting chess board printed via the \\pkg{chess} package. At right, the final position for one of the authors' recent wins (a checkmate playing Black)."}
# library(chess)
# # Sys.setenv(RETICULATE_PYTHON = "my_env/bin/python3")
# game() |> plot()
# 
# library(chessR)
# us <- c("fatgreggy")
# greg <- get_raw_chessdotcom(
#   usernames = us,
#   year_month = 202206
# )
# 
# m <- extract_moves_as_game(greg[6, ])
# # plot_moves(m, interactive = FALSE)
# plot(m)
```

>>>>

Every sport is different, with its own set of rules, strategies, methods of data collection, number of players, and the magnitude of the role of chance. 
At the same time, many sports are similar, either because one evolved from the other, or the structure of the game shares certain attributes. 
Sports that are closely related anthropologically may or may not share common applications of analytical methods. 
For example, despite belonging to the same bat-and-ball family, baseball and cricket differ in strategies such as batting order or sacrifice plays.
Conversely, with just a few small tweaks, analytical metrics might work just as well across sports that are unrelated and quite different.
For instance, an Elo rating could be equally valid for chess players and ice hockey teams.

@@

A more in-depth overview of the history of expected points in sports is provided in @Yurko2019nflwar (Section 1.1). 
Most importantly, this paper uses publicly available data provided by the **nflscrapR** package [@R-nflscrapR] to model the expected points on a play-by-play level in football.
The authors introduce a multinomial logistic regression approach, which takes into account the current down, time remaining, yards from endzone, yards to go, and the indicators for goal to go and whether the time remaining in the half is less than two minutes. 
This model estimates the probabilities of the following possible scoring outcomes after each play: no score, safety, field goal, and touchdown for both the offensive and defensive teams, all of which have a point value. 
The expected points for a play can then be calculated accordingly, by summing up the products of the scoring event point values and their associated probabilities.

@@

The notion of EPV has also been extended to association football (soccer.)
@fernandez2021framework implement deep learning methods to examine the instantaneous expected value of soccer possessions. 
This approach considers passes, ball drives, and shots in soccer as the main set of actions used to compute expected possession value. 
Many applications can be derived from this framework, including predicting which footballer on the pitch is most likely to receive the next pass from the current on-ball player.

@@

```{r chess, fig.show="hold", out.width="49%", fig.cap="At left, the starting chess board printed via the \\pkg{chess} package. At right, the final position for one of the authors' recent wins (a checkmate playing Black)."}
library(chess)
# Sys.setenv(RETICULATE_PYTHON = "my_env/bin/python3")
game() |> plot()

library(chessR)
us <- c("fatgreggy")
greg <- get_raw_chessdotcom(
  usernames = us,
  year_month = 202206
)

m <- extract_moves_as_game(greg[6, ])
# plot_moves(m, interactive = FALSE)
plot(m)
```
